Machine Learning Models for the Segmentation of eCoaching Text

### 2018-03-06 paper-amia-2018/manuscript.tex
investigate different RNN models

However, a significant barrier to fully automate eCoaching is the unsegmented text data. Developing an automatic classification of clinical interactions required segmented text. Nevertheless, eCoaching data comprised of email responses which need to be segmented into a group of MI behavior refers to ``block of text''. Automatic segmentation of eCoaching intervention sessions is a challenging task due to the 2 important reasons. First, the email is an unstructured text that contains informal email exchange in non-traditional formats. Second, a text segment not necessarily belongs to the entire sentence or collection of sentences. One sentence can be segmented into several MI behaviors and vice versa. Figure~\ref{fig:text-segment} illustrates the segmentation of an eCoaching email exchange where the first sentence segmented into 2 different MI behaviors. On the other hand, fourth and fifth segments contain only one and multiple sentences, respectively.

@@

Previous studies mainly focus on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence} in the medical domain. Apostolova et al.\cite{apostolova2009automatic} applied SVM by utilizing word-vector cosine similarity metric combined with several heuristics to classify clinical report into semantic sections such as demographics, history, exam procedure, finding, impression, etc. After identification of each line in the document, Tepper et al. \cite{tepper2012statistical} trained Maximum Entropy models for the section classification. In 2009, Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing technique, terminology-based rule, and naive Bayesian score for identifying sections and headers that achieved 99\% recall with 95.6\% precision. On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with posodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary. However, segmentation of clinical text, in particular, segmentation of MI or eCoaching text into a group of MI behavior is ignored while relying on manual hand-coded approach. Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of MI or eCoaching text into ``block of text''.

@@

Our proposed approach is a novel approach for text segmentation where the previous study focuses on the segmentation of text into sections, headers, and sentences. However, this study segmented email exchange into ``block of text'' represent a particular MI behavior. This work will significantly reduce the amount of resource and time required to segment email text manually. Furthermore, this paper can help to annotate each segment automatically by building a new classifier, which will accelerate the pace of finding best communication strategies to develop an effective MI intervention for healthy eating.

@@

\section*{Conclusion}
Segmentation of eCoaching text is an integral part of developing an automated eCoaching intervention. Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into a group of MI behavior in the setting of discourse analysis with email under the principle of motivational interviews. In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text. We found out that k-nearest neighbor provides the best performance for the segmentation of text in terms of all performance metrics. Manual segmentation of e-coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions. Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model. This approach will help for developing fully automated eCoaching and also accelerate the pace of identifying effective communication strategies linked to healthy eating.

>>>>

However, a significant barrier to fully automate eCoaching is the unsegmented text data. Developing an automatic classification of clinical interactions required segmented text. Nevertheless, eCoaching data comprised of email responses which need to be segmented into groups of MI behaviors. Automatic segmentation of eCoaching intervention sessions is a challenging task due to the 2 important reasons. First, the email is an unstructured text that contains informal email exchange in non-traditional formats. Second, a text segment not necessarily belongs to the entire sentence or collection of sentences. One sentence can be segmented into several MI behaviors and vice versa. Figure~\ref{fig:text-segment} illustrates the segmentation of an eCoaching email exchange where the first sentence segmented into 2 different MI behaviors. On the other hand, fourth and fifth segments contain only one and multiple sentences, respectively.

@@

Previous studies mainly focus on segmentation of text into sections and headers\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2016sentence} in the medical domain. Apostolova et al.\cite{apostolova2009automatic} applied SVM by utilizing word-vector cosine similarity metric combined with several heuristics to classify clinical report into semantic sections such as demographics, history, exam procedure, finding, impression, etc. After identification of each line in the document, Tepper et al. \cite{tepper2012statistical} trained Maximum Entropy models for the section classification. In 2009, Denny et al.\cite{denny2009evaluation} proposed a SecTag algorithm, which combined natural language processing technique, terminology-based rule, and naive Bayesian score for identifying sections and headers that achieved 99\% recall with 95.6\% precision. On the other hand, SVM exploiting with linear kernel and recurrent convolutional neural networks with posodic, part of speech features and word embeddings, were trained by Kreuzthaler et al.\cite{kreuzthaler2015detection} and Griffis et al.\cite{griffis2016quantitative}, respectively, for the detection of sentence boundary. However, segmentation of clinical text, in particular, segmentation of MI or eCoaching text into groups of MI behaviors, is ignored while relying on manual hand-coded approach. Therefore, this study introduces a novel approach and the authors are not aware of any other work this approach has been considered for the segmentation of eCoaching text.

@@

Our proposed approach is novel for text segmentation of eCoaching text because previous studies focus on the segmentation of text into sections, headers, and sentences. However, this study segmented email exchange into groups of MI behavior. This work will significantly reduce the amount of resource and time required to segment email text manually. Furthermore, this paper can help to annotate each segment automatically by building a new classifier, which will accelerate the pace of finding best communication strategies to develop an effective MI intervention for healthy eating.

@@

\section*{Conclusion}
Segmentation of eCoaching text is an integral part of developing an automated eCoaching intervention. Although several studies have done for the segmentation of clinical text into sections and sentences, none of them are used for the segmentation of text into a groups of MI behaviors in the setting of discourse analysis with email under the principle of motivational interviews. In this paper, we compared the performance of machine learning models for the task of segmentation of e-coaching text. We found out that k-nearest neighbor provides the best performance for the segmentation of text in terms of all performance metrics. Manual segmentation of e-coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop effective behavioral interventions. Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model. This approach will help for developing fully automated eCoaching and also accelerate the pace of identifying effective communication strategies linked to healthy eating.
